---
title: "MU Door Scanner Analysis"
author: "Bryson Pursell"
date: "11/18/25"
output: md_document
---

**TLDR: Doors would often not unlock after being scanned, sought to discover if the reason was device, individual person, or method of tapping NFC chip. After organizing random sampling and running observations, analysis discovered that these variables do not significantly predict whether the door will unlock or not.**

**Introduction: The concept for this research began when my roommate and I kept having to re-scan our Miami ID in order to enter our dorm room, when the process is supposed to involve only one quick tap. After speaking to other students around campus, we found that many people were having the same problems, and so I decided to run a statistical analysis to further examine this issue. Upon further discussion with a professor, I landed upon a Binary Response Generalized Linear Model. And thus, the project was off!**


**Beginning Research: To start, I wanted to make sure an experiment of this type was reasonable. To do so, I reached out to Jonathan James, who is an Honors Advisor here at Miami. We had a good conversation about the idea, and he directed me to Angeline Poling, an IT worker at Miami who was versed in the tap card readers used to scan into the rooms. After discussion, I learned that the school uses CBORD CS Access for the software end of this process. For hardware, the SCHLAGE aptiQ MT15 Multi-Technology Single Gang Reader. These card readers have a continous powering feature that ensures the device does not go into a short-term sleep mode. Now, onto the analysis.**

**Design of Experiments: When figuring out the DOE, the final research question was as follows: Is there a relationship between unlock method, device type, and successful door unlocks? For this research, our null hypothesis was that the method of unlocking and number of successful door unlocks have no relationship. The alternative hypothesis was that the method of unlocking and number of successful unlocks do have a relationship. The experimental unit for the experiment was the one door to our dorm (did not have the facilities to access other doors), and as such device + person was limited to those that could access our door (they are coded to individuals). The three factors for such an experiment are method of unlocking the door (Flat on the surface, top of device perpendicular to surface, "Cattycornered" / only corner of device touching scanner, and hovering near the door), device used to unlock (iPhone 13 Pro, Samsung Galaxy S23 Ultra, and Apple Watch 10), and the people unlocking the door (Myself, Logan (roommate), and Ollie (Friend willing to help)). Given these specifications of 4 method, 3 devices, and 3 people (With people acting as a block variable to account for possible noise from attempt inconsistencies), we have 4x3x3 = 36 treatments. As such, we must seek 10-20 replicates to ensure data is relatively accurate. Again related to facilities, we were not permitted more than 1 hour to gather our data.**

**Running Experiment: Attempting to ensure a randomized design, we randomized the person, device, and method for each observation. Using this logic, we gathered 360 observations, using the 36 treatments replicated 10 times.**

**Final Results: Spoiler Warning! After running the experiment I had collected my own data, but wanted confirmation of validity. Using my contact with Joyce Looby, I obtained the results from the 1 hour of data collection on the back-end. This data came directly from the CS Access system, and confirmed that there are times where the system records a successful unlock but one does not occur. Below, you will find the full analysis conducted in R Studio. In short, the model appeared promising to start, but a later discovery revealed that it is not valid. There still remains an issue with the door unlock system, but it is not directly related to the Person, Device, or method used to unlock the door.**

```{r Setup, include=FALSE}

# Initial libraries loaded
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
library(ggfortify)

# Establishment of variables
People <- c("Bryson", "Ollie", "Logan") ## Individuals scanning the door
Devices <- c("Galaxy", "iPhone", "Watch") ## Devised used to scan the door
Method <- c("Flat", "Perp", "Hover", "Cat") ## Method of tapping door scanner
Num_reps <- 10 ## Number of times each combination of people, devices, and methods is repeated

## Establishing data frame to contain synthetic repetitions of data
df <- bind_rows(replicate(Num_reps, expand.grid(People, Devices, Method), simplify = FALSE ) )
names(df) <- c("Person", "Device", "Method")

## Shuffles data to ensure randomization / reduce bias
set.seed(1)
df_shuffled <- df |>
  mutate(Rand_number = runif(NROW(df))) |>
  arrange(Rand_number) |>
  select(-Rand_number)

```

----

**Initial glimpse of data:**
```{r DataLoad_and_Clean, echo=FALSE}

## Reading in collected data and cleaning
door_data <- read.csv("Door_data.csv")

door_data <- door_data |> ## Setting Person variable as a factor and Open variable as a binary response (1 = success, 0 = failure)
  mutate(Person = factor(Person),
         Open = as.logical(Result))

glimpse(door_data) ## Glimpse of table to evaluate effects of cleaning

```

----

**Summary of Fitted Model:**
```{r Model_and_Summary, echo=FALSE}

# Fitted model using generalized linear model (glm), success/failure contingent upon the predictor variables
full_door <- glm(Open ~ Person + Device + Method, 
                 data = door_data, 
                 family = binomial(link = "logit"))

# Outputting results of model summary
summary(full_door)

```

**Total deviance explained by model is 330.76-234.98 = 95.78 (null - residual). This value is large enough that we shall proceed with analysis.**

----

**Whole Model Test:**
```{r FullAnova, echo=FALSE}

# Fitted glm without predictors to be compared against full model
null_risk_fit <- glm(Open ~ 1,
                     data = door_data,
                     family = binomial(link = "logit"))

# Analysis of variance to compare models
anova(null_risk_fit, full_door)

```

**Chi-square value of 95.779 on 7 degrees of freedom (p-value = 2.2e-16). Model including Method, Device, and Person is significantly different from the one without these variables. Pesudo R-squared (1 - 234.98/330.76)*****100 = 28.96% of the deviance is explained by the model. This deviance result shows us that while the model is significant against an empty model, it might not be the best predictor that can be found.**

----

**Lkelihood Ratio Test (LRT)**
```{r LRT, echo=FALSE}

# Try removing one variable at a time to see if any predictors are insignificant
drop1(full_door, test = "LRT")

```

**Method and Device seem to be significant, but Person does not. It is possible that this being the block variable is having an effect. We must now compare the full model to one without the Person variable to find out.**

----

**Reduced model - Person**
```{r ReducPerson, echo=FALSE}

# New fitted glm
nopeople <- glm(Open ~ Method + Device,
                data = door_data,
                family = binomial(link="logit"))

# Anova between -person and full
anova(nopeople, full_door)

```

**Having a Chi-square of .71325 on 2 degrees of freedom (p-value of .7), we do not have evidence that the inclusion of People in the model has a significant effect on the response.**

----

**Exploratory Data Analysis (EDA)**
```{r eda, message=FALSE, echo=FALSE}

# GGpairs to show all possibly relevant data
ggpairs(door_data)

```

**After discovering that a predictor was not significant, a check-up on other variables is necessary**
**Analyzing this EDA plot reveals that the Samsung Galaxy and Perpendicular method seem to have a larger number of failures than other predictors. This must now be explored.**

----

##### No Perpendicular Model

**Glimpse of data w/o Perp method: **
```{r noPerp, echo=FALSE}

# Reading in secondary data set, having replaced every instance of a "Perp" combination with an empty space 
noperp <- read.csv("Sheet 2-Data w o Perpendiculars.csv")
noperp <- noperp |>
  mutate(Open = Result == "1")

glimpse(noperp)

```

----

**No-Perp Model Summary: **
```{r noPerpSumm, echo=FALSE}

noperpdoor <- glm(Open ~ Person + Device + Method,
                  data = noperp,
                  family = binomial(link = "logit"))
summary(noperpdoor)
```

**The total deviance explained by this model is 152.48 - 145.25 = 7.23. Unfortunately, this finding reveals that with the Perp observations removed, the model is not statistically significant from the null model. Thus, our end result iw failure to reject the null hypothesis.**

----

**ANOVA to Fully Confirm Insignificance: **
```{r noPerpAnova, echo=FALSE}
null_perp <- glm(Open ~ 1,
                     data = noperp,
                     family = binomial(link = "logit"))
anova(null_perp, noperpdoor)
```

**Chi-square value of 7.2346 on 6 degrees of freedom (p-value = .2997 > alpha = .05). This further proves that the model does not predict the response.**

**
**

**While the model did not achieve the original goal of the experiment, it does still highlight that there is an issue of small extent in the system. Further analysis may reveal the cause of this issue, but not with this line of reasoning.**
